Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.

### 2019-02-20 papers/CADE19/paper.tex
added explanation about AWR experiments

\section{Optimising Age/Weight Ratios}
\todo{talk about optimising for fixed age/weight}.\\
\todo{talk about experiments with changing age/weight}.

@@

\todo{analyse (two sets of) results, discuss drop in performance in exchange for uniques}.

>>>>

\section{Optimising Age/Weight Ratios}
\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
A base-2 logarithm is used to produce a continuous scale for the \(x\)-axis.
For sufficiently negative \(x\), proof search is biased towards age to such an extent that a proof is found via breadth-first search (and hence the number of active clauses remains constant), whereas after \(x\) becomes sufficiently positive, proof search does not terminate for this problem in the given time.
In between these limits, the function is ``noisy'' and discontinuous.
These properties are typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

\subsection{Fixed AWR}
Changing Vampire's fixed AWR has a significant effect on the number of clauses required to be processed before a proof is found.
For example, Figure \ref{fig:example-optimal-awr} shows the effect of varying AWR on an illustrative problem from TPTP: a smaller number of active clauses means that fewer clauses were processed, which in general means that a proof was found faster\footnote{
It should be noted that if a small number of clauses are extremely expensive to process it may be slower than a larger number of less-expensive clauses, but in general this is a good heuristic measure for prover performance.
It also avoids reproducibility issues involved with using system timing approaches.
}.
On this problem, a good AWR value is over 400\% better by this metric than the worst AWR value.

This experiment was repeated on the whole TPTP problem set, excluding problems Vampire does not support, and problems which were not solved by the default mode in under a second for any AWR value.
The whole set yielded similar AWR data for 8,123 problems.
These data show that while there is no ``best'' AWR for the whole set (Figure \ref{fig:no-best-awr}), choosing a good AWR value for a problem is well-rewarded (Figure \ref{fig:optimal-awr-improvement}).

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{no-best-awr}
	\caption{The mean relative performance drop (in terms of activations compared to an optimal value) choosing a single AWR over an entire problem set causes. No setting is significantly ``better'' or ``worse'' than the others, but some have more variance in performance.}
	\label{fig:no-best-awr}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{optimal-awr-improvement}
	\caption{The distribution of the relative performance gain choosing the best AWR value creates, relative to the worst AWR value.}
	\label{fig:optimal-awr-improvement}
\end{figure}

\subsection{Variable AWR}
Although choosing a good AWR value is important, this is covered in part by the use of strategy scheduling~\cite{CADE18} in which many AWR values are tried in sequence (along with other prover options).
Additionally, a constant AWR fixed for the entire proof search is unlikely to be optimal for any given problem.
This can be shown by running Vampire with a randomised sequence of age/weight ratios given by a random walk repeatedly, then finding the best after a large number of repetitions.
Applying this method with 10,000 repetitions to the problem seen earlier (\texttt{PRO017+2}) yields the example AWR trend shown in Figure \ref{fig:random-walk}, which reduces the best number of activations from 330 with a fixed AWR, to 287 with a varying AWR.

In experiments, several AWR ``shapes'' were commonly seen, including:
\begin{itemize}
	\item Near-constant, as seen currently.
	\item A decreasing trend from a higher AWR to a lower AWR.
	\item An increasing trend from a lower AWR to a higher AWR.
\end{itemize}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{random-walk}
	\caption{The AWR series that produced the lowest number of activations on a particular problem. This is a search strategy that a single fixed AWR cannot reproduce.}
	\label{fig:random-walk}
\end{figure}

@@

\subsection{Decaying AWR}
\subsection{Converging AWR}
\todo{analyse (two sets of) results, discuss drop in performance in exchange for uniques}.
